Synthesis of Linear Functional Programs

### 2021-05-08 article/article.tex
related work introduction?

\todo[inline]{Os~\cite{DBLP:conf/pldi/OseraZ15,DBLP:conf/popl/FrankleOWZ16}
  como ``estrategia geral'' de ``ler'' um sistema de tipos +
  engenharia como um sistema de sintese de programas tipificados;
  Aqui~\cite{DBLP:conf/pldi/PolikarpovaKS16} estamos num contexto
  ``mais avançado'' (polimorfismo e tipos refinados -- ``logica de
  primeira ordem'') -- tipos ricos -- trabalho aspiracional,
  tratamento de funcao recursivas, tratamento de tipos de dados
  indutivos/algebricos; Aqui~\cite{DBLP:conf/lopstr/HughesO20},
  sintese para graded modal types, tem a ver com ``linearidade'' kind
  of, abordagem baseado em focusing (portanto Curry-Howard++), no caso
  deles acaba por ser uma lógica ``modal'' vs linear -- não permite
  futuramente explorar relações com e.g. concorrência, análise de
  recursos.}

Concrete works

\mypara{Type-and-Example-Directed Program Synthesis}~\cite{DBLP:conf/pldi/OseraZ15,DBLP:conf/popl/FrankleOWZ16}
This paper explores a purely functional, recursive program synthesis technique based on types/proof search (as we intend to do),
with examples as an auxiliar technique to trim down the program synthesis vast search space.
\todo[inline]{more introduction to the paper}.
A good use is made of the general strategy of using a type system
to generate programs (by ``inverting'' it), rather than to type-check one.
Moreover, the paper uses a data structure called ``refinement tree'' in conjunction
with the extra information provided by the examples to allow for efficient synthesis
of non-trivial programs.

\item Type-and-example-directed
  synthesis~
  or how to turn your type system upside down.

  \item Synthesis for graded types (no recursion   -- in that work, no
  obvious connection of grading with
  concurrency)~\cite{DBLP:conf/lopstr/HughesO20}

\item Synthesis from (polymorphic) refinement types~\cite{DBLP:conf/pldi/PolikarpovaKS16}

>>>>

% \todo[inline]{Os~\cite{DBLP:conf/pldi/OseraZ15,DBLP:conf/popl/FrankleOWZ16}
%   como ``estrategia geral'' de ``ler'' um sistema de tipos +
%   engenharia como um sistema de sintese de programas tipificados;
%   Aqui~\cite{DBLP:conf/pldi/PolikarpovaKS16} estamos num contexto
%   ``mais avançado'' (polimorfismo e tipos refinados -- ``logica de
%   primeira ordem'') -- tipos ricos -- trabalho aspiracional,
%   tratamento de funcao recursivas, tratamento de tipos de dados
%   indutivos/algebricos; Aqui~\cite{DBLP:conf/lopstr/HughesO20},
%   sintese para graded modal types, tem a ver com ``linearidade'' kind
%   of, abordagem baseado em focusing (portanto Curry-Howard++), no caso
%   deles acaba por ser uma lógica ``modal'' vs linear -- não permite
%   futuramente explorar relações com e.g. concorrência, análise de
%   recursos.}

Type-based program synthesis is a studied field\todo{ dizer de uma forma melhor? }, and it follows
that a lot of literature is available to inspire and complement our work.
Most? follow the synthesis-as-proof-search pattern, however, the process is
novel for each, due to a variety of different rich types explored, and their respective logics; or nuances done to the process, such
as considering examples; or even the programming paradigm of the output produced (e.g. generating heap manipulating programs~\cite{DBLP:journals/pacmpl/PolikarpovaS19}).
Some other common patterns we might want to follow are the type refinements and the support for recursive functions.
These works share some common ground, and base themselves of the same works we do -- most famously the Curry-Howard correspondence, but also,
for example, works on the method of Focusing~\cite{focusing}.

\mypara{Type-and-Example-Directed Program Synthesis}~\cite{DBLP:conf/pldi/OseraZ15,DBLP:conf/popl/FrankleOWZ16}
This paper explores a purely functional, recursive program synthesis technique based on types/proof search (as we intend to do),
with examples as an auxiliar technique to trim down the program synthesis vast search space.
\todo[inline]{more introduction to the paper}.
A good use is made of the general strategy of using a type system
to generate programs (by ``inverting'' the type system), rather than using it to type-check one.
Moreover, the paper uses a data structure called ``refinement tree'' in conjunction
with the extra information provided by the examples to allow for efficient synthesis
of non-trivial programs. Overall it employs good engineering to achieve type-based synthesis
of functional typed programs, with at least two additions that we might want to follow too
(type refinements and recursive functions).

% or how to turn your type system upside down ~ não apagar a Kubrick reference? :)

\mypara{Synthesis from (polymorphic) refinement types}~\cite{DBLP:conf/pldi/PolikarpovaKS16}

\mypara{Synthesis for graded types} (no recursion   -- in that work, no
  obvious connection of grading with
  concurrency)~\cite{DBLP:conf/lopstr/HughesO20}
